Non-kissing and non-crossing complexes \\ for locally gentle algebras

### 2018-07-11 draft/accordionComplexSurfaces.tex
typos

\usepackage[T1]{fontenc}
\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, mathrsfs, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap, xargs, multicol, pdflscape, multirow, hvfloat, array, ae, aecompl, pifont, mathtools, a4wide, float}
\usepackage{marginnote}
\hypersetup{colorlinks=true, citecolor=darkblue, linkcolor=darkblue}
\usepackage[all]{xy}
%\usepackage[bottom]{footmisc}
\usepackage{tikz}
\usepackage{tkz-graph}
\usetikzlibrary{trees, decorations, decorations.markings, shapes, arrows, matrix, calc, fit, intersections, patterns, angles}
\graphicspath{{figures/}}
\makeatletter\def\input@path{{figures/}}\makeatother
\usepackage{caption}
\captionsetup{width=\textwidth}

@@

% COMBINATORICS
% quiver
\newcommand{\blossom}{^\text{\ding{96}}} % blossom
%\newcommand{\blinkers}[1]{_{\LEFTCIRCLE \!\! #1 \!\! \RIGHTCIRCLE}} % blossom
%\newcommand{\indeg}{\mathrm{indeg}} % in-degree
%\newcommand{\outdeg}{\mathrm{outdeg}} % out-degree
%\newcommand{\tai}{\mathsf{tai}} % target incompatible
%\newcommand{\tac}{\mathsf{tac}} % target compatible
%\newcommand{\soi}{\mathsf{soi}} % source incompatible
%\newcommand{\soc}{\mathsf{soc}} % source compatible
%\newcommand{\cso}{\mathsf{cso}} % cosource
%\newcommand{\cta}{\mathsf{cta}} % cotarget
\newcommand{\Enrs}[1]{E_{nr}^{s}(#1)}
\newcommand{\Ers}[1]{E_{r}^{s}(#1)}
\newcommand{\Enrt}[1]{E_{nr}^{t}(#1)}
\newcommand{\Ert}[1]{E_{r}^{t}(#1)}

% walks
%\newcommand{\strings}{\mathcal{S}} % strings
%\newcommand{\distinguishableStrings}{\mathcal{S}_\mathrm{dist}} % distinguishable strings
%\newcommand{\lfStrings}{\underline{\mathcal{S}}} % loop-free strings
%\newcommand{\walks}{\mathcal{W}} % walks
%\newcommand{\straightWalks}{\mathcal{W}_\mathrm{str}} % straight walks
%\newcommand{\bendingWalks}{\mathcal{W}_\mathrm{bend}} % bending walks
%\newcommand{\NKWalks}{\mathcal{W}_\mathrm{nk}} % non-kissing walks
\newcommand{\peaks}[1]{\mathsf{peaks}(#1)} % peaks
\newcommand{\deeps}[1]{\mathsf{deeps}(#1)} % deeps
\newcommand{\distinguishedWalk}[2]{\mathsf{dw}(#1,#2)} % distinguished walk
\newcommand{\distinguishedArrows}[2]{\mathsf{da}(#1,#2)} % distinguished arrows
\newcommand{\distinguishedString}[2]{\mathsf{ds}(#1,#2)} % distinguished string
\newcommand{\distinguishedSign}[2]{\varepsilon(#1,#2)} % distinguished sign

@@

% accordion and slalom complexes
\newcommand{\surface}{\mathcal{S}} % an orientable surface
\newcommand{\dual}{^*} % duality
\newcommand{\dissection}{\mathrm{D}} % dissection of a marked surface
\newcommand{\vertices}{\mathcal{V}} % vertices of a dissection
\newcommand{\edges}{\mathcal{E}} % edges of a dissection
\newcommand{\faces}{\mathcal{F}} % faces of a dissection
\renewcommandx{\AC}[1][1=\dissection]{\mathcal{K}_{\mathrm{acc}}(#1)} % accordion complex
\newcommandx{\RAC}[1][1=\dissection]{\mathcal{C}_{\mathrm{acc}}(#1)} % reduced accordion complex
\newcommandx{\SC}[1][1=\dissection\dual]{\mathcal{K}_{\mathrm{sla}}(#1)} % slalom complex
\newcommandx{\RSC}[1][1=\dissection\dual]{\mathcal{C}_{\mathrm{sla}}(#1)} % reduced slalom complex
\newcommandx{\NCC}[1][1={\dissection, \dissection\dual}]{\mathcal{K}_{\mathrm{nc}}(#1)} % non-crossing complex
\newcommandx{\RNCC}[1][1={\dissection, \dissection\dual}]{\mathcal{C}_{\mathrm{nc}}(#1)} % non-crossing complex
\newcommand{\curveof}{\operatorname{\gamma}} % curve (of a walk)
\newcommand{\edgeof}{\operatorname{\varepsilon}} % arc of a vertex
\newcommand{\dualedgeof}{\operatorname{\varepsilon}\dual} % dual arc of a vertex
\newcommand{\SSS}{\reflectbox{$\mathsf{Z}$}} % S
\newcommand{\ZZZ}{\mathsf{Z}} % Z
\newcommand{\vnext}{n} % next point on the boundary of the surface
\newcommand{\vprevious}{p} % previous point on the boundary of the surface

@@

\begin{theorem*}[\ref{prop:purity} and \ref{coro:thin}]
 The complexes~$\NKC$ and~$\NCC$ are pure and thin.  Their dimension is computed, and mutation is explicitly described.
\end{theorem*}

We end in Section \ref{sec:vectors} with a discussion of~$\b{g}$-vectors,~$\b{c}$-vectors and~$\b{d}$-vectors associated to walks and curves.
Consequences on the representation theory of locally gentle algebras will be investigated in a future project.

Finally, let us briefly review the algebraic and geometric objects which appear in this paper.

@@

The idea of associating a finite-dimensional algebra to a dissection (or a triangulation) of a surface seems to take its roots in the theory of cluster algebras and cluster categories.
This was first done for triangulations of polygons in \cite{CalderoChapotonSchiffler}, and then for any orientable surface with boundary in \cite{ABCP, Labardini}.  
The algebra of a dissection as we shall use it in this paper has appeared in \cite{DavidRoeslerSchiffler}.
In most of the above cases, the algebras obtained are gentle algebras.
It has also been shown in \cite{BaurSimoes} that any gentle algebra is obtained from a dissection of a surface, and that the module category of the algebra can be interpreted by using curves on the surface.
In the case where the surface is a polygon, the $\tau$-tilting theory of the algebra of a dissection has been studied in \cite{PaluPilaudPlamondon, PilaudPlamondonStella}.

Conversely, the construction of a surface associated to a gentle algebra has appeared in \cite{OpperPlamondonSchroll}.
We give a different construction of the same surface in this paper, which is obtained by ``glueing'' quadrilaterals to the arrows of the blossoming quiver (as defined in \cite{PaluPilaudPlamondon}, and called ``framed quiver'' in \cite{BrustleDouvilleMousavandThomasYildirim}).
Our construction has the advantage that it easily yields the two dual dissections of the surface at the same time (the dissection and dual lamination of \cite{OpperPlamondonSchroll}).
We note that our dissections are always cellular, while those found in \cite{BaurSimoes} can be arbitrary.
Remarkably, gentle algebras and surfaces were linked recently in \cite{HaidenKatzarkovKontsevich}, where the Fukaya category of the surface is shown to be equivalent to the bounded derived category of the associated gentle algebra.

@@

\begin{definition}
A \defn{string} for~$\bar Q$ is a word which is either a finite string or an eventually cyclic infinite string.
%We denote by~$\strings(\bar Q)$ the set of strings on~$\bar Q$.
We often implicitly identify the two inverse strings~$\rho$ and~$\rho^{-1}$, and call it an \defn{undirected string} of~$\bar Q$.
%we let~$\strings^\pm(\bar Q) \eqdef \set{\{\rho, \rho^{-1}\}}{\rho \in \strings(\bar Q)}$ denote the collection of all \defn{undirected strings} of~$\bar Q$.
\end{definition}

@@

\begin{definition}
\label{def:walk}
A \defn{walk} of a locally gentle quiver~$\bar Q$ is a maximal string of its blossoming quiver~$\bar Q \blossom$ (meaning that at each end it either reaches a blossom vertex of~$\bar Q \blossom$ or enters an infinite oriented cycle).
%We denote by~$\walks(\bar Q)$ the set of all walks on~$\bar Q$.
As for strings, we implicitly identify the two inverse walks~$\omega$ and~$\omega^{-1}$, and call it an \defn{undirected walk} of~$\bar Q$.
%we let~$\walks^\pm(\bar Q) \eqdef \set{\{\omega, \omega^{-1}\}}{\omega \in \walks(\bar Q)}$ denote the set of \defn{undirected walks} of~$\bar Q$.
\end{definition}

\begin{definition}
\label{def:substrings}
A \defn{substring} of a walk~$\omega = \prod_{i < \ell < j} \alpha_\ell^{\varepsilon_\ell}$ of~$\bar Q$ is a string~$\sigma = \prod_{i' < \ell < j'} \alpha_\ell^{\varepsilon_\ell}$ of~$\bar Q$ for some~$i \le i' < j' \le j$, where the inequality~$i \le i'$ (resp.~$j' \le j$) is strict when~$i \ne -\infty$ (resp.~$j \ne \infty$). In other words, $\sigma$ is a factor of~$\omega$ such that
\begin{itemize}
\item the endpoints of~$\sigma$ are not allowed to be the possible blossom endpoints of~$\omega$,
\item the position of~$\sigma$ as a factor of~$\omega$ matters (the same string at a different position is considered a different substring).
\end{itemize}
Note that the string~$\varepsilon_v$ is a substring of~$\omega$ for each occurence of~$v$ as a vertex of~$\omega$.
We denote by~$\Sigma(\omega)$ the set of substrings of~$\omega$.
We use the same notation for undirected walks (of course, substrings of an undirected walk are undirected).
\end{definition}

@@

\begin{definition}
\label{def:straightBending}
A \defn{peak} (resp.~\defn{deep}) of a walk~$\omega$ is a substring of~$\omega$ of length zero which is on top (resp.~at the bottom of~$\omega$).
A \defn{corner} is either a peak or a deep (in other words, it is a vertex of~$\omega$ where the arrows change direction).
A walk~$\omega$ is \defn{straight} if it has no corner (\ie if~$\omega$ or~$\omega^{-1}$ is a path in~$\bar Q\blossom$), and \defn{bending} otherwise.
%Let $\straightWalks(\bar Q)$ and~$\bendingWalks(\bar Q)$ denote the sets of straight and bending walks of~$\bar Q$ and let~$\straightWalks^\pm(\bar Q) \eqdef \set{\{\omega, \omega^{-1}\}}{\omega \in \straightWalks(\bar Q)}$ and~$\bendingWalks^\pm(\bar Q) \eqdef \set{\{\omega, \omega^{-1}\}}{\omega \in \bendingWalks(\bar Q)}$.
A \defn{peak walk} (resp.~\defn{deep walk}) is a walk with a unique corner (in other words, it switches orientation only once), which is a peak (resp.~deep). For~$v \in Q_0$, we denote by~$v_\peak$ the peak walk with peak at~$v$ and by~$v_\deep$ the deep walk with deep~at~$v$.
\end{definition}

@@

\begin{definition}%[\cite{McConville, PaluPilaudPlamondon}]
\label{def:kiss}
Let~$\omega$ and~$\omega'$ be two undirected walks on~$\bar Q$.
We say that~$\omega$ \defn{kisses}~$\omega'$ if ${\Sigma_\top(\omega) \cap \Sigma_\bottom(\omega')}$ contains a \emph{finite} substring.
See \fref{fig:kissing}.
% \ie if there exist a common substring~$\sigma \in \strings^\pm(\bar Q)$ of~$\omega$ and~$\omega'$ such that the arrows of~$\omega$ incident to~$\sigma$ are both outgoing while the arrows of~$\omega'$ incident to~$\sigma$ are both incoming.
We say that~$\omega$ and~$\omega'$ are \defn{kissing} if~$\omega$ kisses~$\omega'$ or~$\omega'$ kisses~$\omega$ (or both).
Note that we authorize the situation where the commmon finite substring is reduced to a vertex~$v$, meaning that~$v$ is a peak of~$\omega$ and a deep of~$\omega'$.
Observe also that~$\omega$ can kiss~$\omega'$ several times, that~$\omega$ and~$\omega'$ can mutually kiss, and that~$\omega$ can kiss itself.
%We denote by~$\NKWalks(\bar Q)$ the walks on~$\bar Q$ which are not self-kissing and by~$\NKWalks^\pm(\bar Q) \eqdef \set{\{\omega, \omega^{-1}\}}{\omega \in \NKWalks(\bar Q)}$ their undirected version.
%
\begin{figure}[h]
	\capstart
	\centerline{\includegraphics[scale=1]{kissing}}
	\caption{A schematic representation of two kissing walks: $\omega$~kisses~$\omega'$.}
	\label{fig:kissing}
\end{figure}
%
\end{definition}

\begin{definition}%[\cite{McConville, PaluPilaudPlamondon}]
\label{def:nKc}
The \defn{non-kissing complex} of~$\bar Q$ is the simplicial complex~$\NKC$ whose faces are the collections of pairwise non-kissing walks of~$\bar Q$.
Note that self-kissing walks never appear in~$\NKC$ by definition.
In contrast, no straight walk can kiss another walk by definition, so that they appear in all facets of~$\NKC$.
The \defn{reduced non-kissing complex}~$\RNKC$ is the simplicial complex whose faces are the collections of pairwise non-kissing bending walks of~$\bar Q$.
\end{definition}

@@

\begin{figure}[h]
	\capstart
	\centerline{\includegraphics[scale=.7]{halfStar} \qquad \includegraphics[scale=.7]{star}}
	\caption{Two possible situations around~$v$: on the left,~$v$ corresponds to a finite maximal path of~$\bar Q$, while on the right,~$v$ corresponds to an infinite one.}
	\label{fig:stars}
\end{figure}

@@

\begin{remark}
\label{rem:propertiesSurface}
The following observations are useful for the computation of examples.
\begin{enumerate}[(i)]
\item The set~$V_{\bar Q}$ has one vertex for each straight walk in~$\bar Q$ (equivalently, for each maximal path in~$\bar Q$).
      Finite straight walks yield vertices on the boundary, while infinite cyclic straight walks in~$\bar Q$ and in~$\bar Q\koszul$ yield punctures.
\item \label{item:arcs}
      The dissection~$\dissection_{\bar Q}$ has one edge for each vertex~$p \in \bar Q_0$, obtained by concataining the sides~$\Enrt{\alpha} = \Enrs{\beta}$ and~$\Enrt{\alpha'} = \Enrs{\beta'}$ where~$p = t(\alpha) = s(\beta) = t(\alpha') = s(\beta')$, $\alpha\beta \notin I$ and~$\alpha'\beta' \notin I$. We denote by~$\edgeof(p)$ the edge of~$\dissection_{\bar Q}$ corresponding to~$p$.
\item The dissection~$\dissection_{\bar Q}$ has one $\ell$-cell for each straight walk of length~$\ell$ in~$\bar Q\koszul$.
\item A similar remark holds dually for~${V_{\bar Q}}\dual$ and~${\dissection_{\bar Q}}\dual$.
\item The number~$p$ of punctures of~$\surface_{\bar Q}$ is the number of infinite straight walks in~$\bar Q$ and in~$\bar Q\koszul$.
\item \label{item:boundary}
      The number~$b$ of boundary components of~$\surface_{\bar Q}$ can be computed as follows.
      There are two natural perfect matchings whose vertices are the blossom vertices of~$\bar Q$: one is obtained by joining the endpoints of each finite straight walk of~$\bar Q$, and the other is obtained similarly from~$\bar Q \koszul$.
      Let~$G$ be the superposition of these two perfect matchings.
      Then the number~$b$ of boundary components of~$\surface_{\bar Q}$ is the number of connected components of~$G$.
\item The genus of the surface~$\surface_{\bar Q}$ is
      \[
       g = \frac{|Q_1| - |Q_0| - b - p + 2}{2},
      \]
      where~$b$ is the number of boundary components (see above for a way to compute~$b$) and~$p$ the number of punctures (\ie infinite straight walks in~$\bar Q$ and in~$\bar Q\koszul$).
      Indeed, $2g = 2-\chi(\surface_{\bar Q})$, where~$\chi(\surface_{\bar Q})$ is the Euler characteristic of~$\surface_{\bar Q}$.
      The dissection~$\dissection$ provides a cellular decomposition of~$\surface_{\bar Q}$ on which we fill the boundary components and the punctures with disks.  
      The number of faces of this cellular decomposition is~$b+p$, its number of edges is~$|Q_0|$ and its number of vertices is the number of finite straight walks in~$\bar Q$, which is equal to~$2|Q_0| - |Q_1|$.
      The above formula for~$g$ follows.
\end{enumerate}
\end{remark}

@@

We apply Remark~\ref{rem:propertiesSurface} to understand the corresponding surfaces.
The corresponding perfect matchings (see Remark~\ref{rem:propertiesSurface}\,(\ref{item:boundary})) form a cycle with green and red arrows alternating.
Moreover, $\bar Q$ and~$\bar Q\koszul$ have no infinite straight walks.
Since~$|Q_1| = |Q_0|-1$, the corresponding surfaces have $1$ boundary component, no puncture and genus~$0$.
Two examples are represented in \fref{fig:reversedPathsSurfaces}.
Note that the dissection~$\dissection_{\bar Q}$ is a triangulation with no internal triangles, and the orientation of~$\bar Q$ indicates how to glue these triangles.
For instance, the Cambrian quiver completely oriented in one direction yields a fan triangulation, were all internal arcs are incident to the same vertex, as in \fref{fig:reversedPathsSurfaces}\,(left).

@@

\begin{figure}[H]
	\capstart
	\centerline{\includegraphics[scale=.45]{doubleCyclesQuivers}}
	\caption{The double cycle quivers with~$1, 2, 3, 4, 5$ vertices.}
	\label{fig:doubleCyclesQuivers}
\end{figure}

>>>>

\usepackage[T1]{fontenc}
\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, mathrsfs, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap, xargs, multicol, pdflscape, multirow, hvfloat, array, ae, aecompl, pifont, mathtools, a4wide}
\usepackage{marginnote}
\hypersetup{colorlinks=true, citecolor=darkblue, linkcolor=darkblue}
\usepackage[all]{xy}
%\usepackage[bottom]{footmisc}
\usepackage{tikz}
\usepackage{tkz-graph}
\usetikzlibrary{trees, decorations, decorations.markings, shapes, arrows, matrix, calc, fit, intersections, patterns, angles}
\graphicspath{{figures/}}
\makeatletter\def\input@path{{figures/}}\makeatother
\usepackage{caption}
\captionsetup{width=\textwidth}

@@

% COMBINATORICS
% quiver
\newcommand{\blossom}{^\text{\ding{96}}} % blossom
\newcommand{\blinkers}[1]{_{\LEFTCIRCLE \!\! #1 \!\! \RIGHTCIRCLE}} % blossom
\newcommand{\indeg}{\mathrm{indeg}} % in-degree
\newcommand{\outdeg}{\mathrm{outdeg}} % out-degree
%\newcommand{\tai}{\mathsf{tai}} % target incompatible
%\newcommand{\tac}{\mathsf{tac}} % target compatible
%\newcommand{\soi}{\mathsf{soi}} % source incompatible
%\newcommand{\soc}{\mathsf{soc}} % source compatible
%\newcommand{\cso}{\mathsf{cso}} % cosource
%\newcommand{\cta}{\mathsf{cta}} % cotarget
\newcommand{\Enrs}[1]{E_{nr}^{s}(#1)}
\newcommand{\Ers}[1]{E_{r}^{s}(#1)}
\newcommand{\Enrt}[1]{E_{nr}^{t}(#1)}
\newcommand{\Ert}[1]{E_{r}^{t}(#1)}

% walks
\newcommand{\strings}{\mathcal{S}} % strings
\newcommand{\distinguishableStrings}{\mathcal{S}_\mathrm{dist}} % distinguishable strings
\newcommand{\lfStrings}{\underline{\mathcal{S}}} % loop-free strings
\newcommand{\walks}{\mathcal{W}} % walks
\newcommand{\straightWalks}{\mathcal{W}_\mathrm{str}} % straight walks
\newcommand{\bendingWalks}{\mathcal{W}_\mathrm{bend}} % bending walks
\newcommand{\NKWalks}{\mathcal{W}_\mathrm{nk}} % non-kissing walks
\newcommand{\peaks}[1]{\mathsf{peaks}(#1)} % peaks
\newcommand{\deeps}[1]{\mathsf{deeps}(#1)} % deeps
\newcommand{\distinguishedWalk}[2]{\mathsf{dw}(#1,#2)} % distinguished walk
\newcommand{\distinguishedArrows}[2]{\mathsf{da}(#1,#2)} % distinguished arrows
\newcommand{\distinguishedString}[2]{\mathsf{ds}(#1,#2)} % distinguished string
\newcommand{\distinguishedSign}[2]{\varepsilon(#1,#2)} % distinguished sign

@@

% accordion and slalom complexes
\newcommand{\surface}{\mathcal{S}} % an orientable surface
\newcommand{\dual}{^*} % duality
\newcommand{\dissection}{\mathrm{D}} % dissection of a marked surface
\newcommand{\vertices}{\mathcal{V}} % vertices of a dissection
\newcommand{\edges}{\mathcal{E}} % edges of a dissection
\newcommand{\faces}{\mathcal{F}} % faces of a dissection
\renewcommandx{\AC}[1][1=\dissection]{\mathcal{K}_{\mathrm{acc}}(#1)} % accordion complex
\newcommandx{\RAC}[1][1=\dissection]{\mathcal{C}_{\mathrm{acc}}(#1)} % reduced accordion complex
\newcommandx{\SC}[1][1=\dissection\dual]{\mathcal{K}_{\mathrm{sla}}(#1)} % slalom complex
\newcommandx{\RSC}[1][1=\dissection\dual]{\mathcal{C}_{\mathrm{sla}}(#1)} % reduced slalom complex
\newcommandx{\NCC}[1][1={\dissection, \dissection\dual}]{\mathcal{K}_{\mathrm{nc}}(#1)} % non-crossing complex
\newcommandx{\RNCC}[1][1={\dissection, \dissection\dual}]{\mathcal{C}_{\mathrm{nc}}(#1)} % non-crossing complex
\newcommand{\curveof}{\operatorname{\gamma}} % curve (of a walk)
\newcommand{\arcof}{\operatorname{\varepsilon}} % arc of a vertex
\newcommand{\dualarcof}{\operatorname{\varepsilon}\dual} % dual arc of a vertex
\newcommand{\SSS}{\reflectbox{$\mathsf{Z}$}} % S
\newcommand{\ZZZ}{\mathsf{Z}} % Z

@@

\begin{theorem*}[\ref{prop:purity} and \ref{coro:thin}]
 The complexes~$\NKC$ and~$\NCC$ are pure and thin.  Their dimension is computed, and mutation is explicitly described.
\end{theorem*}
We end in Section \ref{sec:vectors} with a discussion of~$\b{g}$-vectors,~$\b{c}$-vectors and~$\b{d}$-vectors associated to walks and curves.

Consequences on the representation theory of locally gentle algebras will be investigated in a future project.

Let us review the algebraic and geometric objects which appear in this paper.

@@

The idea of associating a finite-dimensional algebra to a dissection (or a triangulation) of a surface seems to take its roots in the theory of cluster algebras and cluster categories.
This was first done for triangulations of polygons in \cite{CalderoChapotonSchiffler}, and then for any orientable surface with boundary in \cite{ABCP, Labardini}.  
The algebra of a dissection as we shall use it in this paper has appeared in \cite{DavidRoeslerSchiffler}.
In most of the above cases, the algebras obtained are gentle algebras.
It has also been shown in \cite{BaurSimoes} that any gentle algebra is obtained from a dissection of a surface, and that the module category of the algebra can be interpreted by using curves on the surface.
In the case where the surface is a polygon, the $\tau$-tilting theory of the algebra of a dissection has been studied in \cite{PaluPilaudPlamondon, PilaudPlamondonStella}.
\vincent{Should we also mention cluster algebras from surfaces in this paragraph?}
\pierreguy{Not sure.}

Conversely, the construction of a surface associated to a gentle algebra has appeared in \cite{OpperPlamondonSchroll}.
We give a different construction of the same surface in this paper, which is obtained by ``glueing'' quadrilaterals to the arrows of the blossoming quiver (as defined in \cite{PaluPilaudPlamondon}, and called ``framed quiver'' in \cite{BrustleDouvilleMousavandThomasYildirim}).
Our construction has the advantage that it easily yields the two dual dissections of the surface at the same time (the dissection and dual lamination of \cite{OpperPlamondonSchroll}).
We note that our dissections are always cellularr, while those found in \cite{BaurSimoes} can be arbitrary.
Remarkably, gentle algebras and surfaces were linked recently in \cite{HaidenKatzarkovKontsevich}, where the Fukaya category of the surface is shown to be equivalent to the bounded derived category of the associated gentle algebra.

@@

\begin{definition}
A \defn{string} for~$\bar Q$ is a word which is either a finite string or an eventually cyclic infinite string.
We denote by~$\strings(\bar Q)$ the set of strings on~$\bar Q$.
We often implicitly identify the two inverse strings~$\rho$ and~$\rho^{-1}$, and we let~$\strings^\pm(\bar Q) \eqdef \set{\{\rho, \rho^{-1}\}}{\rho \in \strings(\bar Q)}$ denote the collection of all \defn{undirected strings} of~$\bar Q$.
\vincent{Can we delete this notation.}
\end{definition}

@@

\begin{definition}
\label{def:walk}
A \defn{walk} of a locally gentle quiver~$\bar Q$ is a maximal string of its blossoming quiver~$\bar Q \blossom$ (meaning that at each end it either reaches a blossom vertex of~$\bar Q \blossom$ or enters an infinite oriented cycle).
We denote by~$\walks(\bar Q)$ the set of all walks on~$\bar Q$.
As for strings, we implicitly identify the two inverse walks~$\omega$ and~$\omega^{-1}$, and we let~$\walks^\pm(\bar Q) \eqdef \set{\{\omega, \omega^{-1}\}}{\omega \in \walks(\bar Q)}$ denote the set of \defn{undirected walks} of~$\bar Q$.
\vincent{Can we delete this notation.}
\end{definition}

\begin{definition}
\label{def:substrings}
A \defn{substring} of a walk~$\omega = \prod_{i < \ell < j} \alpha_\ell^{\varepsilon_\ell} \in \walks(\bar Q)$ is a string~$\sigma = \prod_{i' < \ell < j'} \alpha_\ell^{\varepsilon_\ell}$ of~$\strings(\bar Q)$ for some~$i \le i' < j' \le j$, where the inequality~$i \le i'$ (resp.~$j' \le j$) is strict when~$i \ne -\infty$ (resp.~$j \ne \infty$). In other words, $\sigma$ is a factor of~$\omega$ such that
\begin{itemize}
\item the endpoints of~$\sigma$ are not allowed to be the possible blossom endpoints of~$\omega$,
\item the position of~$\sigma$ as a factor of~$\omega$ matters (the same string at a different position is considered a different substring).
\end{itemize}
Note that the string~$\varepsilon_v$ is a substring of~$\omega$ for each occurence of~$v$ as a vertex of~$\omega$.
We denote by~$\Sigma(\omega)$ the set of substrings of~$\omega$.
We use the same notation for undirected walks (of course, substrings of an undirected walk are undirected).
\end{definition}

@@

\begin{definition}
\label{def:straightBended}
A \defn{peak} (resp.~\defn{deep}) of a walk~$\omega$ is a substring of~$\omega$ of length zero which is on top (resp.~at the bottom of~$\omega$).
A \defn{corner} is either a peak or a deep (in other words, it is a vertex of~$\omega$ where the arrows change direction).
A walk~$\omega$ is \defn{straight} if it has no corner (\ie if~$\omega$ or~$\omega^{-1}$ is a path in~$\bar Q\blossom$), and \defn{bending} otherwise.
%Let $\straightWalks(\bar Q)$ and~$\bendingWalks(\bar Q)$ denote the sets of straight and bending walks of~$\bar Q$ and let~$\straightWalks^\pm(\bar Q) \eqdef \set{\{\omega, \omega^{-1}\}}{\omega \in \straightWalks(\bar Q)}$ and~$\bendingWalks^\pm(\bar Q) \eqdef \set{\{\omega, \omega^{-1}\}}{\omega \in \bendingWalks(\bar Q)}$.
A \defn{peak walk} (resp.~\defn{deep walk}) is a walk with a unique corner (in other words, it switches orientation only once), which is a peak (resp.~deep). For~$v \in Q_0$, we denote by~$v_\peak$ the peak walk with peak at~$v$ and by~$v_\deep$ the deep walk with deep~at~$v$.
\end{definition}

@@

\begin{definition}%[\cite{McConville, PaluPilaudPlamondon}]
\label{def:kiss}
Let~$\omega,\omega' \in \walks^\pm(\bar Q)$ be two undirected walks on~$\bar Q$.
We say that~$\omega$ \defn{kisses}~$\omega'$ if~$\Sigma_\top(\omega) \cap \Sigma_\bottom(\omega')$ contains a \emph{finite} substring.
See \fref{fig:kissing}.
% \ie if there exist a common substring~$\sigma \in \strings^\pm(\bar Q)$ of~$\omega$ and~$\omega'$ such that the arrows of~$\omega$ incident to~$\sigma$ are both outgoing while the arrows of~$\omega'$ incident to~$\sigma$ are both incoming.
We say that~$\omega$ and~$\omega'$ are \defn{kissing} if~$\omega$ kisses~$\omega'$ or~$\omega'$ kisses~$\omega$ (or both).
Note that we authorize the situation where the commmon finite substring is reduced to a vertex~$v$, meaning that~$v$ is a peak of~$\omega$ and a deep of~$\omega'$.
Observe also that~$\omega$ can kiss~$\omega'$ several times, that~$\omega$ and~$\omega'$ can mutually kiss, and that~$\omega$ can kiss itself.
We denote by~$\NKWalks(\bar Q)$ the walks on~$\bar Q$ which are not self-kissing and by~$\NKWalks^\pm(\bar Q) \eqdef \set{\{\omega, \omega^{-1}\}}{\omega \in \NKWalks(\bar Q)}$ their undirected version.
\vincent{Can we delete this notation.}
%
\begin{figure}[h]
	\capstart
	\centerline{\includegraphics[scale=1]{kissing}}
	\caption{A schematic representation of two kissing walks: $\omega$~kisses~$\omega'$.}
	\label{fig:kissing}
\end{figure}
%
\end{definition}

\begin{definition}%[\cite{McConville, PaluPilaudPlamondon}]
\label{def:nKc}
The \defn{non-kissing complex} of~$\bar Q$ is the simplicial complex~$\NKC$ whose faces are the collections of pairwise non-kissing walks of~$\NKWalks^\pm(\bar Q)$.
Note that self-kissing walks never appear in~$\NKC$ by definition.
In contrast, no straight walk can kiss another walk by definition, so that they appear in all facets of~$\NKC$.
The \defn{reduced non-kissing complex}~$\RNKC$ is the simplicial complex whose faces are the collections of pairwise non-kissing bending walks of~$\bar Q$.
\end{definition}

@@

\begin{figure}[h]
	\capstart
	\centerline{\includegraphics[scale=.7]{halfStar} \qquad \includegraphics[scale=.7]{star}}
	\caption{The situation around~$v$ (green vertex). On the left,~$v$ corresponds to a finite maximal path of~$\bar Q$, while on the right,~$v$ corresponds to an infinite one.}
	\label{fig:stars}
\end{figure}

@@

\begin{remark}
\label{rem:propertiesSurface}
The following observations are useful for the computation of examples.
\begin{enumerate}[(i)]
\item The set~$V_{\bar Q}$ has one vertex for each straight walk in~$\bar Q$ (equivalently, for each maximal path in~$\bar Q$).
      Finite straight walks yield vertices on the boundary, while infinite cyclic straight walks in~$\bar Q$ and in~$\bar Q\koszul$ yield punctures.
\item \label{item:arcs}
      The dissection~$\dissection_{\bar Q}$ has one arc for each vertex~$p \in \bar Q_0$, obtained by concataining the sides~$\Enrt{\alpha} = \Enrs{\beta}$ and~$\Enrt{\alpha'} = \Enrs{\beta'}$ where~$p = t(\alpha) = s(\beta) = t(\alpha') = s(\beta')$, $\alpha\beta \notin I$ and~$\alpha'\beta' \notin I$.
\item The dissection~$\dissection_{\bar Q}$ has one $\ell$-cell for each straight walk of length~$\ell$ in~$\bar Q\koszul$.
\item A similar remark holds dually for~${V_{\bar Q}}\dual$ and~${\dissection_{\bar Q}}\dual$.
\item The number~$p$ of punctures of~$\surface_{\bar Q}$ is the number of infinite straight walks in~$\bar Q$ and in~$\bar Q\koszul$.
\item \label{item:boundary}
      The number~$b$ of boundary components of~$\surface_{\bar Q}$ can be computed as follows.
      There are two natural perfect matchings whose vertices are the blossom vertices of~$\bar Q$: one is obtained by joining the endpoints of each finite straight walk of~$\bar Q$, and the other is obtained similarly from~$\bar Q \koszul$.
      Let~$G$ be the superposition of these two perfect matchings.
      Then the number~$b$ of boundary components of~$\surface_{\bar Q}$ is the number of connected components of~$G$.
\item The genus of the surface~$\surface_{\bar Q}$ is
      \[
       g = \frac{|Q_1| - |Q_0| - b - p + 2}{2},
      \]
      where~$b$ is the number of boundary components (see above for a way to compute~$b$) and~$p$ the number of punctures (\ie infinite straight walks in~$\bar Q$ and in~$\bar Q\koszul$).
      Indeed, $2g = 2-\chi(\surface_{\bar Q})$, where~$\chi(\surface_{\bar Q})$ is the Euler characteristic of~$\surface_{\bar Q}$.
      The dissection~$\dissection$ provides a cellular decomposition of~$\surface_{\bar Q}$ on which we fill the boundary components and the punctures with disks.  
      The number of faces of this cellular decomposition is~$b+p$, its number of edges is~$|Q_0|$ and its number of vertices is the number of finite straight walks in~$\bar Q$, which is equal to~$2|Q_0| - |Q_1|$.
      The above formula for~$g$ follows.
\end{enumerate}
\end{remark}

@@

We now apply Remark~\ref{rem:propertiesSurface} to understand the corresponding surfaces.
The corresponding perfect matchings (see Remark~\ref{rem:propertiesSurface}\,(\ref{item:boundary})) form a cycle with green and red arrows alternating.
Moreover, $\bar Q$ and~$\bar Q\koszul$ have no infinite straight walks.
Since~$|Q_1| = |Q_0|-1$, the corresponding surfaces have $1$ boundary component, no puncture and genus~$0$.
Two examples are represented in \fref{fig:reversedPathsSurfaces}.
Note that the dissection~$\dissection_{\bar Q}$ is a triangulation with no internal triangles, and the orientation of~$\bar Q$ indicates how to glue these triangles.
For instance, the Cambrian quiver completely oriented in one direction yields a fan triangulation, were all internal arcs are incident to the same vertex, as in \fref{fig:reversedPathsSurfaces}\,(left).

@@

\begin{figure}[h]
	\capstart
	\centerline{\includegraphics[scale=.45]{doubleCyclesQuivers}}
	\caption{The double cycle quivers with~$1, 2, 3, 4, 5$ vertices.}
	\label{fig:doubleCyclesQuivers}
\end{figure}
